---
title: "Introduction to Git and GitHub"
title-slide-attributes: 
  data-background-color: "#43464B"
subtitle: "Session - Workflow with {usethis} and {gert}"
author: "Zoë Turner"
format:
  revealjs:
    theme: [default, nhsr-quarto.scss]
    logo: https://raw.githubusercontent.com/nhs-r-community/assets/main/logo/nhsr-logo.svg
    code-link: true
    css: nhsr-quarto.scss
    preview-links: true
execute: 
  echo: true
  eval: false
embed-resources: true  
---

## Not the command line!

-   Many courses in Git refer to the Command line
-   This is under the `Terminal` tab in RStudio
-   There are many resources available for Git commands but perhaps the hardest part is the workflow
-   Which is where R packages can make things a little friendlier...

## Getting started

Using two packages:

```{r}
library(usethis)
library(gert)
```

::: {.callout-tip collapse="false" appearance="default" icon="true"}
## Terminal or Console?

The `Console` starts with `>` and is for `R`\
The `Terminal` starts with `$` and is for `git` but other things too like `quarto`
:::

## Set up a project

::: {.callout-tip collapse="false" appearance="default" icon="true"}
## Remember!

-   On Posit Cloud the workspace *is* a project!
:::

-   [Setting up a project on your own computer](https://nhs-r-community.github.io/intro-r-rstudio/session-projects.html#/title-slide)
-   The Files should have a file called `*.Rproj` if it's a project

## Start new work

Rather than working on the `main` branch we need a copy called "new_work"

```{r}
usethis::pr_init("new_work")
```

1.  Next [create a quarto file](https://intro-r-rstudio.nhsrcommunity.com/session-quarto.html#/open-a-new-quarto-file)
2.  include a title and author
3.  Render it (which will require you to also save it)

## Commit the new files

In the Git pane three new items will have appeared

-   the `*.qmd` file
-   an `*.html` file from the rendered quarto
-   a folder for the corresponding html (because this isn't self-contained)

## Making a quarto html self-contained

Add to the YAML (the first few lines at the top)

```{r}
#| code-line-numbers: "5"
title: "My report"
author: "Zoë Turner"
format: html
editor: visual
embed-resources: true
```

And render again - the folder will disappear.

## Staging the new files

Individual new files can be `staged` with code or by ticking the box next to the file in the RStudio Git pane

```{r}
gert:git_add("name_of_file.qmd")
```

## Adding multiple files - first method

-   Using [RStudio](https://nhsrcommunity.com/rstudio-and-git-selecting-many-files/) can be done in the Git pane by clicking on the first file name (highlighting it)
-   Then use `Shift` and click the second file to highlight down (or arrow down)
-   Select one of the tick boxes and they all get selected

## Commit

This records the changes with a message in Git

```{r}
gert::git_commit(message = "First commit for quatro document")
```

## Adding multiple files - second method

Stage everything and do the commit message at the same time!

```{r}
gert::git_commit_all(message = "First commit for quatro document")
```

::: {.callout-warning collapse="false" appearance="default" icon="true"}
## Watch out!

This will only `stage` files that have already been "seen" as in committed.

New files will get missed by this command.
:::

## Oops mistake in the message!

Let's correct the misspelling of `quatro` to `quarto` using a feature of RStudio

::: incremental
-   Click on the `Commit` button in the Git pane of RStudio (a pop up will appear)
-   Select the box `Amend previous commit` under the Commit message (starts off blank)
-   Correct `quarot` for `quarto` and then click the Commit button.
:::

## Work in progress

Committing frequently is beneficial but can create a lot of history (commits)

::: incremental
-   A recommended way of working can be to commit with a message "WIP" and update the one commit
-   RStudio can be used to both change the message and add/delete files
:::

## Make a change as WIP

-   Change the title of the quarto document and save
-   Note the appearance of the file in the Git pane with an `M` icon next to it
-   Stage the file using `gert::git_add("name_of_file.qmd")` or tick in RStudio
-   Commit using the RStudio button and select `Amend previous commit`

## Look at the history

-   In the Git pane of RStudio click on `History`
-   Show Diff will need to be selected as the files are new
-   Two links will appear for the `*.qmd` and `*.html`, click on the qmd file
-   You will see your most recent Title name and there is no reference to that because the commit was changed

## A scenario for use

::: incremental
-   In building this course and the slides the content has changed a lot!
-   Slides were merged and split out into topics but the content wasn't always changed
-   WIP on a branch shared on GitHub can show you are still working on this as it's good practice to never have a WIP commit on `main`
:::

## Your turn!

-   Remove the author of the quarto document, save and Render
-   Create a new Quarto presentation and save
-   Use `gert::git_commit_all("Removed author and created slides")`
-   Stage and commit either as a new commit or amend the last

## End session

## Acknowledgements

[Happy Git and GitHub](https://happygitwithr.com/) for the useR by Jenny Bryan

[Pull Request Flow with usethis](https://www.garrickadenbuie.com/blog/pull-request-flow-usethis/?interactive=) by Garrick Aden-Buie
